### 2021-08-02 paper-acm.tex
Initial commit

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%%
%% IMPORTANT NOTICE:
%%
%% For the copyright see the source file.
%%
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%%
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%%
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf,10pt]{acmart}

\settopmatter{printfolios=true}

\usepackage{hyperref}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{A Unified Interaction Model For Web Scraping \& Customization}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.

\author{Kapaya Katongo}
\affiliation{%
  \institution{MIT CSAIL}
  \city{Cambridge, MA}
  \country{USA}
}
\email{kkatongo@mit.edu}

\author{Geoffrey Litt}
\affiliation{%
  \institution{MIT CSAIL}
  \city{Cambridge, MA}
  \country{USA}
}
\email{glitt@mit.edu}

\author{Kathryn Jin}
\affiliation{%
  \institution{MIT CSAIL}
  \city{Cambridge, MA}
  \country{USA}
}
\email{kjin@mit.edu}

\author{Daniel Jackson}
\affiliation{%
  \institution{MIT CSAIL}
  \city{Cambridge, MA}
  \country{USA}
}
\email{dnj@csail.mit.edu}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
% \renewcommand{\shortauthors}{Trovato and Tobin, et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  Web scraping, the process of extracting structured data from a
  website, is a common building block of web customization systems.
  Prior approaches have allowed users to perform web scraping by
  directly demonstrating examples of the data to scrape. However,
  demonstrations alone do not allow for as much expressiveness as
  traditional programming which is vital for complex web scraping tasks.

  In this paper, we present a new interaction model for web scraping
  that combines the ease of use of programming-by-demonstration and the
  expressiveness of traditional programming. When a user demonstrates
  examples of data to extract, a web scraping program is synthesized and
  presented in a formula language with pure functional semantics.
  Crucially, the user can directly edit and execute formulas, allowing
  them to express scraping operations which can not be achieved via
  demonstration alone.

  To illustrate our model, we implement it as a browser extension called
  Joker. Through concrete examples, a small user study and a Cognitive
  Dimenions of Notation analysis, we show how Joker offers an expressive
  interaction model for web scraping and customization.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
%% From HERE
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011066.10011069</concept_id>
<concept_desc>Software and its engineering~Integrated and visual development environments</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Integrated and visual development environments}
% To HERE

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{end-user programming, software customization, web scraping, programming-by-example, program synthesis}

%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
%\begin{teaserfigure}
%  \includegraphics[width=\textwidth]{sampleteaser}
%  \caption{Seattle Mariners at Spring Training, 2010.}
%  \Description{Enjoying the baseball game from the third-base
%  seats. Ichiro Suzuki preparing to bat.}
%  \label{fig:teaser}
%\end{teaserfigure}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\hypertarget{sec:introduction}{%
\section{Introduction}\label{sec:introduction}}

Many websites on the internet do not meet the exact needs of all of
their users. Because of this, millions of people use browser extensions
like Greasemonkey \citep{zotero-224} and Tampermoneky \citep{2021f} to
install userscripts, snippets of Javascript code which customize the
behavior of websites. To make the creation of web customizations more
accessible to end-users without knowledge of programming, researchers
have developed systems like Sifter \citep{huynh2006}, Vegemite
\citep{lin2009} and Wildcard \citep{litt2020, litt2020b}.

A common building block of these web customization systems is web
scraping, the extraction of structured data from websites. They achieve
web scraping in one of two ways: programming-by-demonstration
\citep{huynh2006, lin2009} and traditional programming
\citep{litt2020, litt2020b}. Web scraping by demonstration automates the
web scraping process by utilizing program synthesis to synthesize a web
scraping program from user demonstrations. This approach is accessible
to end-users with no programming experience and enables them to fully
participate in the customization lifecycle. However, web scraping by
demonstration is not powerful enough to perform certain complex web
scraping tasks. On the other side, web scraping by programming involves
programmers manually writing web scraping code. It is powerful enough to
scrape a wider variety of websites but is not accessible to end-users
who can only customize websites that programmers have written web
scraping code for.

In this paper, we present a new interaction model for web scraping for
customization that combines the ease of use of
programming-by-demonstration and the expressiveness of traditional
programming. At its core lies a simple formula language. The language
offers pure functional semantics for expressing and executing data
extractions from websites using Cascading Style Sheet (CSS) selectors.
When a user demonstrates examples of data to scrape, a web scraping
program is synthesized via wrapper induction \citep{kushmerick2000} and
is presented as a formula containing the CSS selector required to
identify the Document Object Model (DOM) element with the data
(Section~\ref{sec:implementation}). Crucially, the formula can be
directly edited and executed to specify complex scraping operations
which can not be achieved by demonstrations alone.

To illustrate our new interaction model for web scraping, we have
implemented it as an extension of Wildcard \citep{litt2020, litt2020b}
called Joker. Wildcard is a browser extension that enables web
customization by direct manipulation of a spreadsheet-like table it adds
to websites. The table contains the website's underlying structured data
and is bidirectionally synchronized with it. User interactions with the
table, including sorting, adding columns and running computations in a
spreadsheet formula language, are propagated to the website thereby
customizing it. The web scraping required to create Wildcard's table can
only be achieved via traditional programming in Javascript. This means
that end-users can only customize websites that programmers have written
web scraping code for. Joker enables end-users to achieve the web
scraping tasks required for customization while maintaining some of the
expressiveness of traditional programming.

By representing web scraping programs as formulas, our web scraping
interaction model fit right into Wildcard's customization paradigm which
utilizes formulas for web customization. This enabled us to go beyond
providing an interaction model that combines web scraping by
demonstration with web scraping by programming to providing an
interaction model that combines web scraping \emph{and} customization.
We refer to this as a \emph{unified interaction model for web scraping
and customization}. In Section~\ref{sec:examples}, we show how this
offers more expressive web scraping and customization.

Our contributions are as follows:

\begin{itemize}
\tightlist
\item
  A \textbf{unified interaction model for web scraping and
  customization} that combines web scraping (by demonstration and
  programing) with web customization through a shared formula language
  with pure functional semantics
\item
  An implementation of the unified interaction model for web scraping
  and customization called Joker that combines three key design
  principles (\textbf{prodirect manipulation}, \textbf{functional
  reactive programming} \& \textbf{unified user interaction}) in a novel
  way (Section~\ref{sec:design-principles})
\item
  The results of an evaluation of the model via an example gallery, a
  five person user study and a Cognitive Dimensions of Notion analysis
  (Section~\ref{sec:evaluation})
\end{itemize}

We end by discussing opportunities for future work and how we can make
CSS selectors more accessible to end-users without development
experience (Section~\ref{sec:conclusion}).

\hypertarget{sec:examples}{%
\section{Example Usage Scenario}\label{sec:examples}}

To concretely illustrate the user experience of our tool, we present a
scenario of customizing eBay, a popular online marketplace. The goal of
the user, Jen, is to filter out search results that are sponsored,
i.e.~product listings that a seller paid to promote. On the webpage,
sponsored listings are marked with a ``Sponsored'' label that Jen can
try to scrape, then she can sort the scraped labels to move all
sponsored listings to the bottom of the page. Figure~\ref{fig:ebay}
shows accompanying screenshots.

\begin{figure*}
  \includegraphics[width=\textwidth]{media/ebay.png}
  \caption{\label{fig:ebay}Scraping and customizing eBay by unified demonstration and formulas.}
\end{figure*}

\emph{Scraping Names by Demonstration} (Figure~\ref{fig:ebay} Part A):
Jen starts the adapter creation process by clicking a context menu item
within the eBay.com page, then hovering over the data that she would
like to scrape: the page element that contains the name of the listing.
The system provides live feedback as Jen hovers: it annotates the row of
data with a border, highlights the column of data in the page with a
green background, and displays how the values will appear in the data
table. When she clicks, the highlighted data is saved into the table.

The text content of the highlighted element is now displayed in the
first column of the table, with one cell for each listing containing the
name of the listing. When she clicks on a cell, Jen can see that the
textual data is also represented by a formula:

\texttt{QuerySelector(rowElement,\ "h3.s-item\_\_title")}

This formula reveals how Joker is actually scraping this column's data
behind the scenes: it calls a query selector on each listing (i.e.~each
\texttt{rowElement}) on the page. Scraped data is stored as functions
that act on the data on the page, instead of as static text, based on
our design principle of \textbf{functional reactive programming}. Joker
exposes this formula to the user to exemplify how scraping is achieved
with formulas and to provide a template for scraping with formulas.

After this first demonstration, Jen has a reference to each of the
page's row elements (i.e.~each product listing) in the Joker table. She
can now use these references to scrape elements within the product
listing with formulas, such as the Sponsored label.

\emph{Scraping Sponsored Labels with Formulas} (Figure~\ref{fig:ebay}
Part B): Next, Jen attempts to scrape contents of the ``Sponsored''
label element into a new column; having this column would allow her to
sort the table based on whether the listing contains the ``Sponsored''
label. She tries to scrape the label using demonstration, but she is
unable to scrape more than one letter of the word at a time. To diagnose
the issue, Jen inspects the page's source code using her browser's
developer tools. Jen discovers that eBay's developers have inserted
invisible letters into word ``Sponsored'' (possibly to obfuscate against
ad blockers). Each letter is in its own HTML element, and the inserted
letters are rendered invisible by CSS. Scraping by demonstration does
not work in this case because Jen wants the element that contains the
whole word, but the system is giving her the leaf-node elements
(individual letters) instead.

Jen sees in the source code that the all of the letters of the
``Sponsored'' label, both visible and invisible, are contained in a
single ancestor element with the CSS selector
\texttt{"div.s-item\_\_title-\/-tagblock"}. Thus, she is able to scrape
the full word by writing a query selector formula with with that CSS
selector. She can copy the formula from the previous column as a
template.

This formula populates the column with the text content of the
``Sponsored'' element, which she can now use for string manipulation and
sorting. By writing a formula, Jen is able to overcome the limitations
of scraping by demonstration. The mixed methods of scraping exemplify
our tool's support of \textbf{prodirect manipulation}.

\emph{Filtering Sponsored Results} (Figure~\ref{fig:ebay} Part C): The
query selector that Jen just wrote returns all of the text within each
of the targeted ``Sponsored'' elements, including any invisible letters.
eBay's web design is that sponsored listings have a visible
``Sponsored'' label with invisible inserted letters, and non-sponsored
listings have an invisible ``Sponsored'' label. Thus, for sponsored
listings, the query returns garbled text
(e.g.~``JSp3onMsoV3rXNYFedZB''), and for non-sponsored listings, the
query returns ``Sponsored.'' Jen identifies this correlation by
scrolling through the scraped data in the column and comparing them to
what she sees on the web page. Then, in a new column, Jen writes a
formula that returns whether or not the previous column's text includes
the word ``Sponsored.'' Finally, she sorts the listings by whether or
not they are sponsored by sorting this column, thus hiding sponsored
results from view. This customization is possible because Joker provides
a \textbf{unified user model}, where scraping and customization are
performed in conjunction.

In this way, Jen is able to use our system to customize the eBay
website, without needing to learn how to program in JavaScript and
without even leaving the webpage. Our model of customization by unified
demonstration and formulas is flexible enough to support a wide range of
other useful modifications and web programming proficiency levels, and
we present a greater variety of use cases in
Section~\ref{sec:evaluation}.

\hypertarget{sec:implementation}{%
\section{System Implementation}\label{sec:implementation}}

In this section, we outline the \emph{wrapper induction}
\citep{kushmerick2000} algorithm that Joker uses to synthesize web
scraping programs from user demonstrations. Then, we briefly describe
the formula language used to represent the synthesized web scraping
programs and include a list of the formulas that are currently available
and their roles.

\hypertarget{wrapper-induction}{%
\subsection{Wrapper Induction}\label{wrapper-induction}}

In order to create web scraping programs from users demonstrations,
Joker solves the wrapper induction \citep{kushmerick2000} task:
generalizing from a few examples of data in a data set to a
specification for all the data in the data set.

Joker takes an approach similar to that used in systems like Vegemite
\citep{lin2009} and Sifter \citep{huynh2006}. It synthesizes a single
\emph{row selector} for the website: a CSS selector that identifies a
set of DOM elements corresponding to the rows of the data set. For each
column in the data set, it synthesizes a \emph{column selector}, a CSS
selector that identifies the element containing the column value.

One important difference is that our algorithm only accepts row elements
that have direct siblings with a similar structure. We refer to this as
the \emph{row-sibling} constraint. Later, we describe how this
constraint provides a useful simplification of the wrapper induction
task and in \citep{section:evaluation} discuss the resulting limitations
this puts on our system. We proceed to describe how CSS selectors are
synthesized for row and column elements and then explain the criteria
used to determine row elements.

\hypertarget{synthesizing-css-selectors}{%
\subsubsection{Synthesizing CSS
Selectors}\label{synthesizing-css-selectors}}

Joker synthesizes two types of CSS selectors: a single row selector that
selects a set of DOM elements corresponding to the rows of the data set
and a column selector for each column which selects the element
containing the column value within a given row.

For a given row element, its row selector is synthesized using the
following criteria:

\emph{Plausibility}. A selector is set of classes. It is a plausible row
selector if it 1) consists of a subset of the classes on the row element
and 2) consists of a subset of the classes all the row element's
siblings. The second requirement is the \emph{row-sibling} constraint we
mentioned. Notice how it simplifies the problem by eliminating
selectors. It is necessary for the end goal of customization because of
customizations like sorting. If row elements are not siblings, the
website would end up in an invalid state when the DOM elements are
re-organized to match the sorting order.

\emph{Weight}. A selector has a weight equal to the number of classes it
consists of.

\emph{Best}. A selector is the best if it is plausible and there is no
other selector that has a lower weight than it. We favor selectors with
the lowest weight to ensure that only the minimum required classes are
utilized. If there are multiple selectors that are plausible and have
the lowest weight, we only pick one.

For a given column element, its column selector is synthesized using the
following criteria:

\emph{Plausibility}. A selector is a plausible column selector if it 1)
consists of a subset of the classes on the column element and 2) only
selects the give column element when applied on the corresponding row
element.

\emph{Weight}. A selector has a weight equal to the number of classes it
consists of.

\emph{Best}. A selector is the best if it is plausible and there is no
other selector that has a lower weight than it has. As before, we favor
selectors with the lowest weight and only pick one if there are multiple
that fulfill the criteria.

One aspect of future work is saving the list of all selectors that
fulfill the criteria and making them available to users to view and pick
from via the formula. This would be similar to Mayer et el's user
interaction model called \emph{program navigation} \citep{mayer2015}
that gives users the opportunity to navigate all valid, synthesized
programs and pick the best one.

\hypertarget{determining-row-elements}{%
\subsubsection{Determining Row
Elements}\label{determining-row-elements}}

When a user first demonstrates a column value, Joker uses the
demonstration to synthesize a row selector. The row selector will be
used to identify all the row elements in the website. Additionally, a
column selector is synthesized that will be used to identify the element
that contains the column value. During subsequent demonstrations, Joker
simply synthesizes a column selector for the column element that
contains the demonstrated column value. Like similar approaches
\citep{huynh2006, lin2009, chasins2018}, all demonstrations have to be
made from the same row element.

Given a demonstrated column value, row elements are determined using the
following criteria:

\emph{Plausibility}. An element \texttt{R} is a plausible row element if
1) it is within the \texttt{BODY} element of the DOM, 2) it is in the
parent path of the column element \texttt{C} containing the demonstrated
column value \texttt{V} and 3) the CSS selector \texttt{S} of element
\texttt{C} identifies a column element \texttt{C\textquotesingle{}} when
applied to \texttt{R}

\emph{Weight}. A row element \texttt{R} has a weight \texttt{W} equal to
the number of its siblings for which the CSS selector \texttt{S} of
column element \texttt{C} only identifies \texttt{C} when applied to it.
This is an enforcement of the \emph{row-sibling} constraint. We want to
find an \texttt{R} with the most number of siblings with a similar
structure

\emph{Best}. A row element \texttt{R} is the best if it is plausible and
there is no other row element that has a higher weight than it. We favor
row elements with the highest weight to ensure that we end up with a
data set with the highest number of column values corresponding to the
column that \texttt{V} belongs to. If there are multiple plausible row
elements with the highest weight, we pick the one closet to the column
element \texttt{C} in its parent path.

Figure~\ref{fig:algorithm} provides a concrete example of how the above
criteria are applied to determine a row element from the demonstration
of a column value.

\begin{figure*}
  \includegraphics[width=\textwidth]{media/algorithm.png}
  \caption{\label{fig:algorithm} An example of how Joker's wrapper indunction algorithm is used to determine the row element from the demonstration of a column value. The row element is correctly determind to be R2 because it is plausible and has the highest weight.}
\end{figure*}

\hypertarget{web-scraping-formulas}{%
\subsection{Web Scraping Formulas}\label{web-scraping-formulas}}

Joker's formula language is similar to that of visual database query
systems like SIEUFERD \citep{bakke2016} and Airtable \citep{2021f}.
Formulas automatically apply across an entire column of data and
reference other column names instead of values in specific rows. This is
more efficient than users having to copy a formula across a column as in
traditional spreadsheets like Microsoft Excel and Google Sheets. It of
course comes at the cost of not being able to specify a formula for only
a subset of column cells but this hasn't yet come up in our use cases.
The language currently consists of the following formulas:

\hypertarget{queryselectorrowelement-selector}{%
\subsubsection{QuerySelector(rowElement,
selector)}\label{queryselectorrowelement-selector}}

This formula is used to represent the web scraping program synthesized
from demonstrations. \texttt{rowElement} is a special keyword that
references a hidden column containing the DOM elements that correspond
to the rows of the data set. \texttt{selector} is the synthesized CSS
selector that specifies the column element.

\hypertarget{getparentelement}{%
\subsubsection{GetParent(element)}\label{getparentelement}}

This formula is used to traverse the DOM when the data to be scraped is
made of the values of its containing and sibling elements.
Demonstrations alone cannot be used to scrape such data.
\texttt{element} can be a reference to a column containing a
\texttt{QuerySelector} formula (\texttt{GetParent(A)}) or a
\texttt{QuerySelector} formula itself
(\texttt{=GetParent(QuerySelector(...))}).

\hypertarget{getattributeelement-attribute}{%
\subsubsection{GetAttribute(element,
attribute)}\label{getattributeelement-attribute}}

This formula is used to scrape data from DOM attributes. An example of
this are URLs which are available on the \texttt{href} attribute of link
elements. \texttt{element} is as described for \texttt{GetParent} and
\texttt{attribute} is the name of the attribute to scrape
(e.g.~\texttt{GetAttribute(A,\ "href")}).

\hypertarget{sec:design-principles}{%
\section{Design Principles}\label{sec:design-principles}}

Below, we describe three design principles that Joker embodies. We did
not invent these principles but rather combined them in a novel manner
for the domain of web scraping and web customization.

\hypertarget{prodirect-manipulation}{%
\subsection{Prodirect Manipulation}\label{prodirect-manipulation}}

\emph{Prodirect Manipulation} is a term coined by Ravi Chugh in a
position paper \citep{chugh2016a} in which he advocates for ``novel
software systems that tightly couple programmatic and direct
manipulation.'' This principles embodies our goal of an interaction
model for web scraping that combines the ease of use of
programming-by-demonstration and the expressiveness of traditional
programming.

In the eBay example in Section~\ref{sec:examples}, we show how the
``Sponsored'' label of a listing can not be scraped by demonstration
alone. However, through Joker's prodirect manipulation interaction, a
user can either directly edit the synthesized web scraping formula or
author one from scratch to achieve the task. This is significant because
the resulting customization to sort the listings by whether they are
sponsored or not would not otherwise be possible

We can see prodirect manipulation in other programming-by-example
systems. Sketch-N-Sketch \citep{chugh2016} allows users to create an SVG
shape via traditional programming and then switch to modifying its size
or shape via direct manipulation. Wrex \citep{drosos2020} takes examples
of data transforms and generates readable and editable Python code.

\hypertarget{functional-reactive-programming}{%
\subsection{Functional Reactive
Programming}\label{functional-reactive-programming}}

Functional reactive programming (FRP) utilizes pure functions to specify
operations in a reactive programming paradigm. FRP has seen wide
adoption in end-user programming through implementations such as
spreadsheet formula languages (Microsoft Excel \& Google Sheets) and
formula languages for low-code programming environments (Microsoft Power
Fx \citep{2021g}, Google AppSheet \citep{2021h}, Airtable \citep{2021f},
Glide \citep{2021a}, Coda \citep{2021c} \& Gneiss \citep{chang2014}).

Joker's use of FRP makes it easier for users to specify scraping
operations programmatically, thus making our web scraping interaction
model more viable. With traditional programming, the eBay example in
Section~\ref{sec:examples} in which demonstrations are not enough to
scrape a listing's ``Sponsored'' label would be much more complicated.
The user would need to understand programming constructs such as state,
variables, looping and data flow in order to write a program to scrape
the label. In Joker, all a user needs to specify is the CSS selector
that identifies the element containing the label. Joker takes care of
managing state, variables, looping and data flow!

A key limitation of this approach is that users need to understand how
CSS selectors work. Because the formula language utilizes pure
functional semantics, users can iterate on CSS selectors as many times
as they need to without having to worry about side effects. This makes
the authoring of CSS selectors more accessible but more work remains for
us to make this fully accessible to end-users.

\hypertarget{unified-user-interaction}{%
\subsection{Unified User Interaction}\label{unified-user-interaction}}

Prior to this work, web scraping and web customization in customization
systems \citep{huynh2006, lin2009} were divided: web scraping had to be
performed prior to customization in a separate phase. Vegemite
\citep{lin2009}, a system for end-user programming of mashups, reported
findings from its user study in which participants thought that ``it was
confusing to use one technique to create the initial table, and another
technique to add information to a new column.''

Joker represents web scraping and customization using the same formula
language. Because of this, both can be performed in a single phase, with
users being able to interleave the two as desired. This makes the
process of customization more incremental which is vital for complex
tasks. We can see this in the Ebay example in
Section~\ref{sec:examples}: when the user scrapes a listing's
``Sponsored'' label, they observe that non-sponsored listings have an
invisible ``Sponsored'' label while sponsored listings have a visible
``Sponsored'' label that consists of a garbled form ``Sponsored.''
Because of Joker's unified user interaction, the user receives the
results of scraping in the table and can immediately write a
customization formula to validate this hypothesis. Without the unified
interaction, the user would have to scrape all the desired columns to
see all their values before ever getting to notice the pattern and
validate their hypothesis.

\hypertarget{sec:evaluation}{%
\section{Evaluation}\label{sec:evaluation}}

In this section, we evaluate Joker with an example gallery of websites,
a small user study and a Cognitive Dimensions of Notion analysis
\citep{blackwell2001}. The example gallery provides examples of websites
that can be scraped and customized by Joker and known limitations. The
user study provides insights on which aspects of Joker work well and
which don't. The Cognitive Dimensions of Notion analysis provides
thoughts on how Joker's design compares to similar approaches and
interaction standards.

\hypertarget{example-gallery}{%
\subsection{Example Gallery}\label{example-gallery}}

Following a method used to evaluate visualizations through a diverse
gallery of examples \citep{ren2018}, our first evaluation of Joker
provides an example gallery of popular websites on which Joker can be
used for web customization and on which it can not be used. For the
websites on which Joker can be used, we provide the sequence of
interactions needed to achieve the customizations. For the websites on
which Joker cannot be used, we provide an explanation.

\hypertarget{websites-joker-can-be-used-on}{%
\subsubsection{Websites Joker Can Be Used
On}\label{websites-joker-can-be-used-on}}

We have used Joker to achieve a variety of purposes across many popular
websites. For example, we have used Joker to sort search results by
price within the Featured page on Amazon. (Using Amazon's sort by price
feature often returns irrelevant results.) In Amazon's source code, the
price is split into three HTML elements: the dollar sign, the dollar
amount, and the cents amount. A user can only scrape the cents element
by demonstration into column A. However, because the parent element of
the cents element contains all three of the price elements, the user can
scrape the full price using the formula \texttt{GetParent(A)}. Next, the
user can write the formula \texttt{ExtractNumber(B)} to convert the
string into a numeric value. Finally, the user can sort this column by
low-to-high prices. In a similar manner, we have used Joker to scrape
and sort prices and ratings on the product listing pages of Target and
eBay.

We have also found Joker to be useful for filtering based on text
inputs. For example, we have used Joker to filter the titles of a
researcher's publications on their Google Scholar profile. Specifically,
a user can first scrape the titles into column A by demonstration. Then,
the user can write the formula \texttt{Includes(A,\ "compiler")} that
returns whether or not the title contains the keyword ``compiler.''
Finally, the user can sort by this column to get all of the publications
that fit their constraint at the top of the page. We have also used
Joker to filter other text-based directory web pages such as Google
search results and the MIT course catalog, in similar ways.

Additionally, we have used Joker to augment web pages with external
information. For example, Joker can augment Reddit's old user interface,
which has a list of headlines with links to articles and images. A user
can first scrape the headline elements into column A by demonstration.
The user can then extract the link into column B with the formula
\texttt{GetAttribute(A,\ "href")}. Then, the user can write the formula
\texttt{ReadTimeInSeconds(B)} that calls an API that returns the links'
read times. Similarly, the user can write the formula
\texttt{Visited(B)} that returns whether that link has been visited in
the user's browser history. The user can also scrape elements such as
the number of comments and the time of posting and sort by these values.
We have performed similar customizations on websites such as ABC news.

\hypertarget{limitations}{%
\subsubsection{Limitations}\label{limitations}}

Joker's unified model of demonstrations and formulas is most effective
on webpages with data that is presented as many similarly-structured
HTML elements. However, certain websites have designs that make it
difficult for Joker to scrape data. These are some of those designs:

\begin{itemize}
\tightlist
\item
  \emph{Multiple row elements.} The layout of some web pages has
  multiple types of row as siblings that contain different children
  elements. For example, the news aggregator website HackerNews has a
  page design that alternates between rows containing a title and rows
  containing supplementary data (e.g.~number of likes and the time of
  posting). Because Joker only chooses a single row selector, when
  scraping by demonstration, Joker will only select one of the types of
  rows, and elements in the other types of rows will not be able to be
  scraped.
\item
  \emph{Infinite scroll.} Some web pages have an ``infinite scroll''
  feature that adds new entries to the page when a user scrolls to the
  bottom. Joker's table will only contain elements that were rendered
  when the table was first created. Additionally, for websites with many
  elements, such as Facebook, Joker might run out of memory while
  running its wrapper induction algorithm and crash the page.
\item
  \emph{Data hidden behind an interaction.} On some sites, a user must
  click on an element to reveal data corresponding to that entry
  (e.g.~time of posting, the author). However, Joker is restricted to
  scraping what is visible on the page at one point in time.
\end{itemize}

\hypertarget{user-study}{%
\subsection{User Study}\label{user-study}}

Our second evaluation of Joker is a user study which provides insights
on what aspects of Joker work well and what don't. Because of this,
successful completion of tasks does not equal validation of Joker. It
does however inform us about usability issues that limit its
effectiveness.

\hypertarget{participants}{%
\subsubsection{Participants}\label{participants}}

We recruited 5 participants with backgrounds ranging from limited
programming experience to Software Engineers. All participants were
familiar with Microsoft Excel spreadsheets but not all had used Excel
spreadsheet formulas. 4 of the participants had web development
experience with 3 of them having extensive experience. 3 of the
participants had web scraping experience with only 1 having extensive
experience.

\hypertarget{tasks}{%
\subsubsection{Tasks}\label{tasks}}

The participants completed 7 tasks across 2 websites towards the goal of
web customization. The first website was MIT's course catalog and the
second was a listing of iPhones after searching for ``iPhone'' on eBay.
The tasks were as follows:

\emph{MIT Course Catalog.} This set of tasks involved web scraping by
demonstration and the use of web customization formulas: 1a) Scrape
course titles, 1b) Scrapes course prerequisites, 1c) Add a column that
indicates whether a course has a prerequisite \& 1d) Add a column that
indicates whether a course does not have a prerequisite and is offered
in the fall.

\emph{eBay.} This set of tasks involved web scraping by demonstration,
the use of web scraping formulas and the use of web customization
formulas: 2a) Scrape iPhone listing title, 2b) Scrape iPhone listing
price, \& 2c) Create a column that indicates whether an iPhone listing
is sponsored.

\hypertarget{protocol}{%
\subsubsection{Protocol}\label{protocol}}

All participants attempted the MIT course catalog and eBay tasks in the
order we have described. We started each session with a description of
Joker and provided a brief tutorial of its main features on a website
not used for the tasks. There was no time limit for tasks but we
provided hints whenever a participant had exhausted the knowledge
available to them.

\hypertarget{results}{%
\subsubsection{Results}\label{results}}

All participants were able to complete all the tasks with the help of
hints when they got stuck and were unable to make progress on their own.
Again, this does not equal validation. We describe the results along the
following dimensions:

\emph{Web Scraping By Demonstration.} All participants where able to
complete the tasks that only involved scraping by demonstration (1a, 1b
\& 2a) quickly and without any hints. For tasks that involved switching
between demonstrations and writing formulas, there was some confusion
about what column demonstrated values would be scraped into. The active
column is managed by a toolbar at the top of the website but
participants hardly noticed it and assumed all operations had to be
performed on the table.

\emph{Web Scraping Formulas.} Participants were unsure why
demonstrations did not work when completing tasks that required using
web scraping formulas (2b, 2c). As a result, they wasted time attempting
to demonstrate using various means. The participant with no web
development experience had the hardest time as expected but was able to
utilize a series of hints we provided to accomplish 2b. The rest of the
participants were able to accomplish 2b with minimal hints but required
a lot more when completing 2c.

\emph{Web Customization Formulas.} Participants had varying degrees of
trouble when completing the tasks that involved using web customization
formulas (1c, 1d, 2c). Most of the confusion resulted from formula
parameters and return values. The formula bar has an autocomplete
feature that shows the documentation for a given formula. The
documentation consists of the parameters a formula it takes and a brief
description of each parameter. However, this wasn't always sufficient to
enable participants to use formulas correctly.

\hypertarget{discussion}{%
\subsubsection{Discussion}\label{discussion}}

One participant, upon seeing our tutorial on scraping by demonstration,
said ``that's like black magic.'' Tasks that involved customization
formulas took longer than those that involved demonstrations but this
was mostly related to the formula documentation not providing enough
usage information. More work is needed to show users what exactly a
given formula does. When asked how they would accomplish the tasks
without Joker, the participant with limited programming experience
responded with ``I don't think I would know how to do it.'' That Joker
enabled them to achieve such a complex task is very promising!

Participants that were familiar with web development were able to take
advantage of Joker's web scraping formulas. This included the
participants that had limited web scraping experience and suggests that
the pure functional semantics and use of CSS selectors for identifying
elements to scrape makes the task easier. As predicted, participants not
familiar with web development didn't have the required background to
understand how CSS selectors work. This highlights the need to further
explore how web scraping formulas can be made more accessible to
end-users.

All participants had trouble figuring out which column the demonstrated
values would be scraped into, especially when demonstrations where
interleaved with use of formulas. This points to the need for the switch
between demonstrating and using formulas to be more tightly coupled
visually. For web scraping formulas, the most notable issue was the lack
of feedback about why demonstrations were not sufficient to scrape a
value. As a result, users wasted time trying to demonstrate in various
ways instead of thinking about using web scraping formulas to achieve
the task. In order for traditional programming to augment demonstrations
as we propose, users need to recieve some kind of feedback that
demonstrations are not sufficient.

\hypertarget{cognitive-dimensions-analysis}{%
\subsection{Cognitive Dimensions
Analysis}\label{cognitive-dimensions-analysis}}

Our third evaluation of Joker analyzes it using Cognitive Dimensions of
Notation \citep{blackwell2001}, a heuristic evaluation framework that
has been used to evaluate programming languages and visual programming
systems \citep{satyanarayan2014, satyanarayan2014a, ledo2018}. When
contrasting our tool with traditional scraping and other visual tools,
we find particularly meaningful differences along several of the
dimensions:

\emph{Progressive evaluation.} In Joker, a user can see the intermediate
results of their scraping and customization work at any point, and
adjust their future actions accordingly. The table user interface makes
it easy to inspect the intermediate results and notice surprises like
missing values.

In contrast, traditional scraping typically requires editing the code,
manually re-running it, and inspecting the results in an unstructured
textual format, making it harder to progressively evaluate the results.
Also, many end-user scraping tools \citep{chasins2018, lin2009} require
the user to demonstrate all the data extractions they want to perform
before showing any information about how those demonstrations will
generalize across multiple examples.\footnote{This is an instance where
  Joker's limitation of only scraping a single page at a time proves
  beneficial. All the relevant data is already available on the page
  without further network requests, making it possible to support
  progressive evaluation with low latency. Other tools that support
  scraping across multiple pages necessarily require a slower feedback
  loop.}

\emph{Premature commitment.} Many scraping tools require making a
\emph{premature commitment} to a data schema: first, the user extracts a
dataset, and then they perform downstream analysis or customization
using that data. Wildcard suffered from this problem: when writing code
for a scraping adapter, a user would need to try to anticipate all
future customizations and extract the necessary data.

Our current system instead supports extracting data \emph{on demand}.
The user can decide on their desired customizations and data
transformations, and extract data as needed to fulfill those tasks.
There is never a need to eagerly guess what data will be needed in
advance.

We have also borrowed a technique from spreadsheets for avoiding
premature commitment: default naming. New data columns are automatically
assigned a single-letter name, so that the user does not need to
prematurely think of a name before extracting the data. We have not yet
implemented the capability to optionally rename demonstrated columns,
but it would be straightforward to do so, and would provide a way to
offer the benefits of names without requiring a premature commitment.

\emph{Provisionality.} Joker makes it easy to try out a scraping action
without fully committing to it. When the user hovers over any element in
the page, they see a preview of how that data would be entered into the
table, and then they can click if they'd like to proceed. This makes it
feel very fast and lightweight to try scraping different elements on the
page.

\emph{Viscosity.} Some scraping tools have high viscosity: they make it
difficult to change a single part of a scraping specification without
modifying the global structure. For example, in Rousillon
\citep{chasins2018}, changing the desired demonstration for a single
column of a table requires re-demonstrating all of the columns. In
contrast, Joker allows a user to change the specification for a single
column of a table without modifying the others, resulting in a lower
viscosity in response to changes.

\emph{Role-expressiveness.} One dimension we are still exploring in our
tool is role-expressiveness: having different elements of a program
clearly indicate their role to the user. In particular, in our current
design, the visual display of references to DOM elements in the table is
similar to the display of primitive values. In our experience, this can
sometimes make it difficult to understand which parts of the table are
directly interacting with the website, vs.~processing the downstream
results. In the future we could consider adding more visual
differentiation to help users understand the role of different parts of
their customization.

\hypertarget{sec:related-work}{%
\section{Related Work}\label{sec:related-work}}

Joker builds on existing work in end-user web scraping, end-user web
customization and program synthesis by a number of systems.

\hypertarget{end-user-web-scraping}{%
\subsection{End-user Web Scraping}\label{end-user-web-scraping}}

FlashExtract \citep{le2014} is a programming-by-example tool for data
extraction. In addition to demonstrating whole values, it supports
demonstrating substrings of values. Joker only supports this through
formulas. This is not as end-user friendly but allows for a wider range
of operations such as indicating whether demonstrated values contain a
certain value or are greater than or less than a certain value.

Rousillon \citep{chasins2018} is a tool that enables end-users to scrape
distributed, hierarchical web data. It presents the web scraping program
generated by demonstration as an editable, high-level, block-based
language called Helena \citep{2021c}. While Helena can be used to
specify complex web scraping tasks like adding control flow, it does not
present the synthesized web scraping program. This means that users can
only scrape what can be demonstrated. Joker on the other hand displays
the synthesized program as a formula which can be modified to increase
the expressiveness of scraping.

\hypertarget{end-user-web-customization}{%
\subsection{End-user Web
Customization}\label{end-user-web-customization}}

Vegemite \citep{lin2009} is a tool for end-user programming of mashups.
Unlike Joker, its table interface is only populated and can only be
interacted with after all the demonstrations have been provided. This
does not support interleaving of scraping and table operations to
achieve an incremental workflow for users. Mashups are created using
CoScripter \citep{leshed2008} which records operations on the scraped
values in the table for automation tasks. CoScripter provides the
generated automation program as text-based commands, such as ``paste
address into `Walk Score' input,'' which can be edited via ``sloppy
programming'' \citep{lin2009} techniques. However, this editing does not
extend to the synthesized web scraping program which is not displayed
and therefore cannot be modified. This means that users can only scrape
what can be demonstrated.

Sifter \citep{huynh2006} is a tool that augments websites with advanced
sorting and filtering functionality. It attempts to automatically detect
items and fields on the page with a variety of clever heuristics. If
this falls, it gives the user the option of demonstrating to correct the
result. In contrast, Joker is simpler and makes fewer assumptions about
the structure of websites by giving control to the user from the
beginning of the process and displaying the synthesized program which
can be modified. We hypothesize that focusing on a tight feedback loop
rather than automation may support a scraping process that is just as
fast as an automated one, offers more expressive scraping and extends to
a greater variety of websites. However, further user user testing is
required to validate this hypothesis.

\hypertarget{program-synthesis}{%
\subsection{Program Synthesis}\label{program-synthesis}}

FlashProg \citep{mayer2015} is a framework that provides program
navigation and disambiguation for programming-by-example tools like
FlashExtract \citep{le2014} and FlashFill \citep{harris}. The program
viewer provides a high level description of synthesized programs as well
as a way to navigate the list of alternative programs that satisfy the
demonstrations. This is important because demonstrations are an
ambiguous specification for program synthesis \citep{peleg2018}: the set
of synthesized programs for a demonstration can be very large. To
further ensure that the best synthesized program is arrived at,
FlashProg has a disambiguation viewer that asks the user questions in
order to resolve ambiguities in the user's demonstrations. In contrast,
Joker only presents the top-ranked synthesized program which may not be
the best one. Furthermore, the program is presented in is low-level form
as a CSS selector. This is not end-user friendly but allows for more
expressiveness.

\hypertarget{sec:conclusion}{%
\section{Conclusion And Future Work}\label{sec:conclusion}}

In this paper, we presented our unified interaction model for web
scraping and customization through a browser extension called Joker.
This model combines the ease of use of programming-by-demonstration and
the expressiveness of traditional programming. At its core lies a simple
formula language that offers pure functional semantics for expressing
and executing data extractions from websites using Cascading Style Sheet
(CSS) selectors.

The main area of future work involves providing feedback to users about
when demonstrations are not sufficient. Otherwise, users won't know to
switch to formulas to take advantage of their increased expressiveness.
FlashProg \citep{mayer2015} and Wrangler \citep{kandel2011}, whose
interfaces also have tables, offer clues for how to provide feedback
about the execution of programs synthesized from demonstrations. Another
important area involves making the formula language more accessible to
end-users not familiar with CSS selectors. Again, FlashProg offers some
clues through its use of high level descriptions to give meaning to
synthesized programs. One avenue for associating meaning with CSS
selectors could be by extracting semantic web content as seen in systems
like Thresher \citep{hogue2005}.

% \printbibliography

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{references-bibtex.bib}

\end{document}
\endinput
%%
%% End of file `sample-sigconf.tex'.